Auditory Scene Similarity Retrieval and Classification with Relevance-based Quantization of Scattering Features

### 2016-12-16 paper/structureScene.tex
polishing abs and intro

un important task in computational ecoacoustics is to determine the similarity of two audio scenes. This is often done using bag-of-frames models and other representations based on summary statistics. While these successfully characterize static scenes with little variability in sound content, they are unable to accurately differentiate scenes characterized by a few distinct sound events superimposed over a static background. To address this problem, we propose a two-scale representation, which describes a signal at a short scale using scattering transforms while relying on a cluster model to characterize its larger-scale structure. To motivate this, we validate the use of scattering transforms for supervised acoustic scene classification (ASC), illustrating its state-of-the-art performance compared to traditional feature transforms like mel-frequency cepstral coefficients (MFCCs). Furthermore, we illustrate the power of the full cluster-based representation by applying it to an acoustic scene similarity retrieval (ASSR) task, where we again significantly outperform more traditional approaches based on summary statistics of MFCCs.

@@

An important problem in this field is the characterization of auditory scenes for the purposes of similarity estimation. One application of this is acoustic scene similarity retrieval (ASSR), where the scenes judged ``most similar'' to a given query scene are extracted from some larger dataset. This requires the auditory scenes to be represented in some way that captures their distinctive qualities. Often, this has been achieved by bag-of-frames approaches, which attempt to describe the auditory scene recording using summary statistics. Unfortunately, since this only captures the average structure of the scene, the approach often fails when presented scenes which are dynamic and contain rare, distinct sound events that discriminate the scene from others.Indeed, experiments in cognitive psychology suggests that this matches more closely the way that humans and other animals perceive acoustic scenes. \ja{Refs?}

@@

The first question to be answered when confronted with this task is how is the signal to be represented at a low level. In other words, how do we represent an extract of the auditory scene signal over a sub-second interval? A standard tool for this are the mel-frequency cepstral coefficients (MFCCs), originally developed for speech processing \cite{davis-mermelstein}, but which have found wider use in recent years \cite{logan}. \ja{Other MFCC works?} A richer set of features are found in the scattering transform, a generic signal representation which has enjoyed significant success in various audio classification tasks \cite{Anden2014}.

The next question is the choice of a classifier. Here, a class of particularly successful algorithms train a classifier, such as a support vector machine (SVM), to discriminate between short extracts of an auditory scene, and at the classification stage divide the signal into short extracts which are classified separately before pooling the classification results to reach a final decision. This last step is known as late integration and is crucial to the success of these methods, since it allows the classifier to concentrate on the short, distinct sound events that discriminate between scenes. The success of the late integration technique again leads credence to the idea that rare, distinct events in an auditory scene are sufficient to characterize it completely. Combining scattering transforms for low-level representation with late-integrated SVMs used for classifying the scene at a higher level, we obtain a competitive auditory scene classification system, demonstrating the usefulness of both concepts for this task.

Returning to scene retrieval setting, we cannot use supervised models such as SVMs, but we can emulate their selection behavior by quantizing the auditory scene and defining a distance over this quantization. We therefore propose a new model for acoustic scenes, where the signal is represented at sub-second scales by scattering transforms, while larger scales are captured by an unsupervised method that quantizes the scattering coefficients into a number of clusters. We then use these to define a set of distances that can subsequently be used for similarity retrieval.Testing this approach on an scene retrieval task, we obtain significant improvements over traditional bag-of-frames and summary statistics models, applied both to MFCCs and scattering coefficients.

>>>>

An important task in computational ecoacoustics is to determine the similarity of two audio scenes. This is often done using bag-of-frames models and other representations based on summary statistics. While these successfully characterize static scenes with little variability in sound content, they are unable to accurately differentiate scenes characterized by a few distinct sound events superimposed over a static background. To address this problem, we propose a two-scale representation, which describes a signal at a short scale using scattering transforms while relying on a cluster model to characterize its larger-scale structure. To motivate this, we validate the use of scattering transforms for supervised acoustic scene classification (ASC), illustrating its state-of-the-art performance compared to traditional feature transforms like mel-frequency cepstral coefficients (MFCCs). Furthermore, we illustrate the power of the full cluster-based representation by applying it to an acoustic scene similarity retrieval (ASSR) task, where we again significantly outperform more traditional approaches based on summary statistics of MFCCs.

@@

An important problem in this field is the characterization of auditory scenes for the purposes of similarity estimation. One application of this is acoustic scene similarity retrieval (ASSR), where the scenes judged ``most similar'' to a given query scene are extracted from some larger dataset. This requires the auditory scenes to be represented in some way that captures their distinctive qualities. Often, this has been achieved by bag-of-frames approaches \cite{aucouturier2007bag}, which attempt to describe the auditory scene recording using summary statistics. Unfortunately, since this only captures the average structure of the scene, the approach often fails when presented scenes which are dynamic and contain rare, distinct sound events that discriminate the scene from others. Indeed, experiments in cognitive psychology suggests that this matches more closely the way that humans and other animals perceive acoustic scenes. \gl{put some Refs}

@@

Adressing this task requires first to be able to capture meanningful aspects within the signal at a small time scale. In other words, how do we represent an extract of the auditory scene signal over a sub-second interval? A standard tool for this are the mel-frequency cepstral coefficients (MFCCs), originally developed for speech processing \cite{davis-mermelstein}, but which have found wider use in recent years in music information retrieval \cite{logan} and environemental audio processing \cite{aucouturier2007bag}. A richer set of features are found in the scattering transform, a generic signal representation which has enjoyed significant success in various audio classification tasks \cite{Anden2014}.

The next question is the choice of a classifier. Here, a class of particularly successful algorithms train a classifier, such as a support vector machine (SVM) \cite{cristianini2000introduction}, to discriminate between short extracts of an auditory scene, and at the classification stage divide the signal into short extracts which are classified separately before pooling the classification results to reach a final decision. This last step is known as late integration and is crucial to the success of these methods, since it allows the classifier to concentrate on the short, distinct sound events that discriminate between scenes. The success of the late integration technique again leads credence to the idea that rare, distinct events in an auditory scene are sufficient to characterize it almost completely. Combining scattering transforms for low-level representation with late-integrated SVMs used for classifying the scene at a higher level, we obtain a competitive auditory scene classification system, demonstrating the usefulness of both concepts for this task.

Returning to scene retrieval setting, we cannot use supervised models such as SVMs, but we can emulate their selection behavior by quantizing the auditory scene and defining a distance over this quantization. We therefore propose a new model for acoustic scenes, where the signal is represented at sub-second scales by scattering transforms, while larger scales are captured by an unsupervised method that quantizes the scattering coefficients into a given number of clusters. We then use these clusters to define a set of distances that can subsequently be used for similarity retrieval. Testing this approach on an scene retrieval task, we obtain significant improvements over traditional bag-of-frames and summary statistics models, applied both to MFCCs and scattering coefficients.
